---
title: "LaBella_04"
author: "Rachel LaBella"
date: "October 2, 2016"
output: html_document
---

####Book Questions
#####Chapter 6 

15.
a) Pygmy mammoths and continental mammoths have the same femur length.
b) Patients taking phentermine and topiramate lose weight at the same rate as the control patients without the drugs.
c) Patients taking phentermine and topiramate have the same rate of babies born with cleft palats than the patients not taking the drugs.
d) There is no difference in candy consumption when Christmas music is playing or normal music is playing.
e) Male white-collared manakins dance the same amount when a female is present or absent. 

21. 
a) There will be a higher chance of getting a type-II error if your sample size is 60. The larger the sample size, the more power the study has. There is a higher chance of rejecting the null if there is greater power.
b) The 100 participant study has a higher power. 
c) Both of the tests have an equal probability of obtaining a type-I error. The only way to decrease the chance of a type-I error is to lower the alpha to something below 0.05. Doing so just increases your chance of getting a type-II error though. 
d) I think that it should be a one-tailed test. You are testing whether COX-2 selective inhibitors have an effect on cardiac arrest or do not. 

29.
a) The researchers would have a 95% chance that they would not reject any of the true null hypothesis. There is a 5% chance of creating a type-1 error and the result would be a false positive. 
b) 95% are expected to reject the null hypothesis, 5% will result in a type-1 error. 

####Chapter 7
22. 
a) 
```{r, echo = TRUE}
library(binom)
binom.confint(6101, 9821, conf.level = 0.95, methods = "exact")
pbinom(6101, 9821, prob = 0.5, lower.tail = FALSE)
```
b) The 95% confidence interval states that 95% of the time the mean will lie within 0.61 and 0.63 probabiities. It is highly unlikely that the buttered toast will have a 50/50 chance, it lies outside the confidence interval. The p-value is insanely low. 


```{r, echo = TRUE}
library(dplyr)
library(ggplot2)
library(tidyr)
```

####3 - Here’s the exercise we started in lab. Feel free to look back copiously at the lab handout if you’re getting stuck. Remember, for each step, write-out in comments what you want to do, and then follow behing with code.
Now, let’s assume an average population-wide resting heart rate of 80 beats per minute with a standard deviation of 6 BPM.
A given drug speeds people’s heart rates up on average by 5 BPM. What sample size do we need to achieve a power of 0.8?

####3.1) Start up your simulation

Make a simulated data frame to look at the effects of multiple sample sizes: from 1-20, with 500 simulations per sample size, and also multiple SD values, from 3 through 10 (just 3:10, no need for non-integer values). You’re going to want crossing with your intitial data frame of just sample sizes and a vector of sd values to start. Then generate samples from the appropriate random normal distribution.
``` {r, echo = TRUE}

# set a mean and pop SD
m <- 80 + 5
sd_pop <- 3:10

sample_df <- data.frame(samp_size = rep(1:20, 500)) 

sd_vector <- c(3:10)

cross_df <- sample_df %>%
  crossing(sd_vector = sd_vector) %>%
  group_by(samp_size, sd_vector) %>%
  ungroup()

sim_cross_df <- cross_df %>%
  group_by(1:n()) %>%
  mutate(sample_mean = mean(rnorm(samp_size, m, sd_vector))) %>%
  ungroup()

ggplot(data=sim_cross_df, mapping = aes(samp_size, sample_mean)) +
  geom_jitter(alpha = 0.4, size = 3)

```

####3.2) Z!

OK, now that you’ve done that, calculate the results from z-tests. Plot p by sample size, using facet_wrap for different SD values.

``` {r, echo = TRUE}

z_data <- sim_cross_df %>%
  mutate(se_y = (sd_vector)/sqrt(samp_size)) %>%
  mutate(z = (sample_mean-80)/se_y) %>%
  mutate(p = pnorm(abs(z), lower.tail=FALSE))

ggplot(data=z_data, mapping = aes(samp_size, p)) +
  facet_wrap(~sd_vector) +
  geom_jitter() +
  theme_bw()

```


####3.3) P and Power

Now plot power for an alpha of 0.05, but use color for different SD values. Include our threshold power of 0.8.

``` {r, echo = TRUE}
power_df <- z_data %>%
  group_by(samp_size, sd_vector) %>%
  summarize(error_rate = sum(p>0.08)/n()) %>%
  ungroup() %>%
  mutate(power = 1 - error_rate)

ggplot(data = power_df, mapping = aes(samp_size, power, group = sd_vector, color = sd_vector)) +
  geom_point() +
  geom_line() + 
  geom_hline(yintercept = 0.8, linetype = 2, color = "red") +
  theme_bw()


```


####3.4) Many alphas

Last, use crossing again to explore changing alphas from 0.01 to 0.1. Plot power curves with different alphas as different colors, and use faceting to look at different SDs.

``` {r, echo = TRUE}

alpha <- seq(0.01, 0.1, .01)
alpha_df <- z_data %>%
  crossing(alpha = alpha) %>%
  group_by(samp_size, alpha, sd_vector) %>%
  summarize(error_rate = sum(p>alpha)/n()) %>%
  ungroup() %>%
  mutate(power = 1 - error_rate)

ggplot(alpha_df, mapping = aes(samp_size, power, color = factor(alpha))) +
  facet_wrap(~sd_vector) +
  geom_line() +
  geom_point() +
  geom_hline(yintercept = 0.8, linetype = 2, color = "red") +
  theme_bw()
  
```


####3.5) What does it all mean? What do you learn about how alpha and SD affect power?
So the alpha level shows us what percentage we are willing to accept a type I error. So for instance, if we have a .05 alpha, we have a 5% chance that we will get a type I error (a false positive). 

In the graphs, when the standard deviation is low, the sample size does not need to be large to get a power of 1.0. As the standard deviation increases, you need a larger sample size to get a power of 0.8 or higher. 

The bigger the standard deviation is, the harder it will be to NOT get a type 1 or 2 error. 


####3.6) How do you think that changing the effect size would affect power?

If you increase the effect size, there will be an easier time seeing changed in sample sizes when they're smaller. If the sample size is smaller, the power is smaller as well. The larger the sample size, the larger the power and the more significant the data. It would depend on the sample size as to whether the effect size would have a large effect on the power. 


